---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(sf, quietly = TRUE)
set.seed(2022)
```

# nswgeo

<!-- badges: start -->
[![nswgeo status badge](https://cidm-ph.r-universe.dev/badges/nswgeo)](https://cidm-ph.r-universe.dev)
<!-- badges: end -->

A collection of geospatial datasets and map plotting helpers for working
with New South Wales maps. The maps are registered with `{cartographer}` and so
are compatible with `{ggautomap}`.

## Installation

You can install nswgeo like so:

``` r
options(repos = c(
  cidmph = 'https://cidm-ph.r-universe.dev',
  CRAN = 'https://cloud.r-project.org'))
  
install.packages('nswgeo')
```

## Examples

The datasets you use with the package will have one column that contains
a feature name (Local Health District, Local Government Area, or postal
code). You'll probably also have columns you want to use for faceting
and/or colouring points:

```{r example-setup}
library(nswgeo)
library(ggautomap)
library(ggplot2)
library(dplyr, warn.conflicts = FALSE)

data(covid_cases_nsw, package = "nswgeo")
covid_cases_nsw
```

You need to let the package know which column has the feature by setting
the `location` aesthetic. This example has three different columns of locations
for different feature types; your dataset only needs to have one of these.

In general you'll start with `geom_boundaries()` to draw the base map. This geom
needs to be told which `feature_type` you're after (e.g. `"nswgeo.lga"` for LGAs).
All of the summary geoms of `{ggautomap}` can then be used to draw your data.

## Scatter

```{r example-scatter, fig.height = 3, fig.width = 7}
covid_cases_nsw %>%
  ggplot(aes(location = lga)) +
  geom_boundaries(feature_type = "nswgeo.lga") +
  geom_geoscatter(aes(colour = type), sample_type = "random", size = 0.5) +
  coord_sf_zoom(datum = NA) +
  theme_minimal()
```

Points are drawn at random within the boundaries of their location.
This example also uses `coord_sf_zoom()` as a replacement for `coord_sf()` that
automatically crops the map around your data points.


## Insets

To show a zoomed in part of the map as an inset, you can configure an inset and
provide it to each relevant geom. The geoms all support `{ggmapinset}`.

```{r example-scatter-inset, fig.height = 3, fig.width = 7}
inset_cfg <- configure_inset(centre = "Blacktown", radius = 40, units = "km",
                             scale = 7, translation = c(400, -100))

covid_cases_nsw %>%
  ggplot(aes(location = lga)) +
  geom_boundaries(feature_type = "nswgeo.lga", inset = inset_cfg) +
  geom_geoscatter(aes(colour = type), sample_type = "random",
                  size = 0.5, inset = inset_cfg) +
  geom_inset_frame(inset = inset_cfg) +
  coord_sf(datum = NA) +
  theme_minimal()
```

## Packed points

This next example uses `geom_centroids()` to place the points in a
packed circle in the centre of each feature. It also shows how you can
fine-tune the plot with the usual `{ggplot2}` functions.

```{r example-pack, fig.height = 3, fig.width = 7}
inset_cfg <- configure_inset(centre = "Western Sydney", radius = 35, units = "km",
                             scale = 8, translation = c(450, -100))

covid_cases_nsw %>%
  filter(year >= 2021) %>%
  # set the scale of the packing differently for the inset
  mutate(scale_km = if_else(lhd == "Western Sydney", 4, 30)) %>%
  ggplot(aes(location = lhd)) +
  geom_boundaries(feature_type = "nswgeo.lhd", inset = inset_cfg) +
  geom_centroids(aes(colour = type, scale = scale_km), size = 1, inset = inset_cfg) +
  geom_inset_frame(inset = inset_cfg) +
  coord_sf(datum = NA) +
  facet_wrap(vars(year)) +
  # use the centroids stat + a fixed nudge to place the text in the right area
  geom_text(
    aes(after_stat(x), after_stat(y), label = lhd),
    stat = "centroids",
    nudge_y = -1.2,
    size = 3,
    data = ~ slice_head(group_by(.x, lhd, year), n = 1)) +
  scale_x_continuous(expand = expansion(add = c(NA, 2))) +
  labs(x = NULL, y = NULL) +
  theme_minimal()
```

## Choropleths

```{r example-inset, fig.height = 3, fig.width = 7}
inset_cfg <- configure_inset(centre = "Western Sydney", radius = 60, units = "km",
                             scale = 3.5, translation = c(350, 0))

covid_cases_nsw %>%
  ggplot(aes(location = lhd, group = NA)) +
  geom_choropleth(inset = inset_cfg) +
  geom_boundaries(feature_type = "nswgeo.lhd", colour = "black", linewidth = 0.1,
                  outline_colour = "black", inset = inset_cfg) +
  geom_inset_frame(inset = inset_cfg) +
  coord_sf(datum = NA) +
  scale_fill_steps(low = "#e6f9ff", high = "#00394d", n.breaks = 5, na.value = "white") +
  labs(x = NULL, y = NULL, fill = NULL) +
  theme_minimal()
```

## Pies and Australia

The `"nswgeo.states"` map data includes the other states for national summaries.
Both the abbreviated state names (e.g. `"NSW"`) and the full names (e.g.
`"New South Wales"`) are supported in the `location` aesthetic.
This example also shows `geom_pie()`, which aggregates points in each location.

```{r example-national}
national_data <- data.frame(state = sample(c("qld", "nsw", "vic", "sa", "act"),
                                           size = 1000, replace = TRUE,
                                           prob = c(0.2, 0.35, 0.3, 0.1, 0.05)),
                            type = sample(c("A", "B", "C"),
                                          size = 1000, replace = TRUE,
                                          prob = c(0.3, 0.6, 0.1)))
national_data$type[national_data$state == "act"] <- "A"

national_data %>%
  ggplot(aes(location = state)) +
  geom_boundaries(feature_type = "nswgeo.states") +
  geom_pie(aes(fill = type), pie_radius = 1.5) +
  coord_sf(datum = NA) +
  theme_minimal()
```
